---
title: "Getting Started with {nemsqar}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with {nemsqar}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(nemsqar)
```

# Introduction

The `{nemsqar}` package provides an automated and reproducible framework for calculating EMS quality measures defined by the National EMS Quality Alliance (NEMSQA). These measures are widely used by EMS agencies, trauma systems, quality improvement teams, and researchers to evaluate performance, identify gaps in care, and support evidence‑based improvement activities.

This vignette is intended for users who are experienced in EMS, injury epidemiology, or quality improvement but who may be beginners in R or in performing measure calculation using R. The goal is to provide a clear, step‑by‑step introduction to loading data, preparing it for use with `nemsqar`, and running selected NEMSQA measures in a consistent and reproducible manner.

By the end of this vignette, users will understand:

- The structure and purpose of the example datasets included in `nemsqar`
- How to load NEMSQA‑ready tables using `nemsqar`'s built in synthetic data
- How to prepare data for measure calculation
- How to run individual NEMSQA measure functions
- How to interpret outputs generated by the package

The following sections walk through the entire workflow, from loading EMS data to producing standardized performance measures that align with national reporting expectations.

# The Measures

This vignette focuses on three NEMSQA measures implemented in `nemsqar`:

- **Asthma‑01**: Assessment and treatment of patients with suspected asthma  
- **Respiratory‑01**: Measurement of respiratory assessment and intervention  
- **Safety‑01**: Evaluation of medication administration safety  

Each measure requires specific input tables and produces outputs in a standardized format. The examples below demonstrate how to prepare the necessary inputs and execute each measure step by step.

# The Data

Before calculating measures, it is important to understand the datasets included with `nemsqar`. The package provides small, NEMSQA‑ready example tables that represent the expected structure of EMS incident, procedure, medication, and vital sign data. These datasets allow users to practice the full workflow without needing to supply their own EMS records.

## Loading Example Data

The `{nemsqar}` package includes several example datasets that mirror the National EMS Information System (NEMSIS) structure required for NEMSQA measure calculation. These datasets are designed for users who are new to R and allow you to practice the workflow before applying the functions to production EMS data.

To load any example dataset, use the `data()` function. The naming pattern `nemsqar_*_table` identifies each table by its content type (for example, scene, patient, procedures, vitals, medications). To acquire the data needed to run performance estimates for Asthma-01, for example, it is necessary to have:

```{r asthma_01_tables, echo=TRUE, eval=TRUE, results="hide"}
data("nemsqar_patient_scene_table")
data("nemsqar_response_table")
data("nemsqar_situation_table")
data("nemsqar_medications_table")
```

After loading, each object will appear in your R environment as a standard data frame.

## Inspecting Table Structure

Because R is new for many users, it is important to confirm that each dataset loaded correctly and contains the variables expected by the measure calculation functions. A few simple commands help verify structure:

### Review the structure of a loaded dataset
```{r}
# An easy to read quick look at the data using the `glimpse()` function from the famous `dplyr` package
dplyr::glimpse(nemsqar_patient_scene_table)
```

### Take a look at the first few rows of the dataset itself
```{r head_patient_scene_table, eval=TRUE, echo=TRUE}
# An abbreviated look at the actual data tables
head(nemsqar_patient_scene_table, n = 10)
```

These functions show variable names, data types, and the first few rows of the table. This step is critical, because each measure expects specific fields, and incorrect data types (for example, character instead of numeric) will cause functions to fail.

## Understanding Required Inputs
Each NEMSQA measure uses a defined set of input tables. `nemsqar` was designed to accept one single data source (via argument `df`) in case a user has their data structured in this way. Generally, though, this is not recommended. It is best practice to store your data in different tables according to the NEMSIS structure such as having 1:1 e*Patient* and *eScene* data in one table, *eResponse* data in a separate table, *eSituation* data in another, and so on. Although the EMS domain concepts may already be familiar, the R functions (and NEMSQA) require them to be organized in a consistent format. For example:

* **Asthma‑01** uses patient, scene, response, situation, and medication tables.
* **Respiratory‑01** uses patient, scene, response, situation, and vitals tables.
* **Safety‑01** uses patient and response tables.

The following sections demonstrate how to supply these inputs to `nemsqar` and how to run your first measure.