---
title: "Getting Started with {nemsqar}"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with {nemsqar}}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  eval = TRUE,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, echo=FALSE, message=FALSE, results="hide"}
library(nemsqar)
```

# Introduction

The `{nemsqar}` package provides an automated and reproducible framework for calculating EMS quality measures defined by the National EMS Quality Alliance (NEMSQA). These measures are widely used by EMS agencies, trauma systems, quality improvement teams, and researchers to evaluate performance, identify gaps in care, and support evidence‑based improvement activities.

This vignette is intended for users who are experienced in EMS, injury epidemiology, or quality improvement but who may be beginners in R or in performing measure calculation using R. The goal is to provide a clear, step‑by‑step introduction to loading data, preparing it for use with `nemsqar`, and running selected NEMSQA measures in a consistent and reproducible manner.

By the end of this vignette, users will understand:

- The structure and purpose of the example datasets included in `nemsqar`
- How to load NEMSQA‑ready tables using `nemsqar`'s built in synthetic data
- How to prepare data for measure calculation
- How to run individual NEMSQA measure functions
- How to interpret outputs generated by the package

The following sections walk through the entire workflow, from loading EMS data to producing standardized performance measures that align with national reporting expectations.

# The Measures

This vignette focuses on three NEMSQA measures implemented in `nemsqar`:

- **Asthma‑01**: Assessment and treatment of patients with suspected asthma  
- **Respiratory‑01**: Measurement of respiratory assessment and intervention  
- **Safety‑01**: Evaluation of medication administration safety  

Each measure requires specific input tables and produces outputs in a standardized format. The examples below demonstrate how to prepare the necessary inputs and execute each measure step by step.

# The Data

Before calculating measures, it is important to understand the datasets included with `nemsqar`. The package provides small, NEMSQA‑ready example tables that represent the expected structure of EMS incident, procedure, medication, and vital sign data. These datasets allow users to practice the full workflow without needing to supply their own EMS records.

## Loading Example Data

The `{nemsqar}` package includes several example datasets that mirror the National EMS Information System (NEMSIS) structure required for NEMSQA measure calculation. These datasets are designed for users who are new to R and allow you to practice the workflow before applying the functions to production EMS data.

To load any example dataset, use the `data()` function. The naming pattern `nemsqar_*_table` identifies each table by its content type (for example, scene, patient, procedures, vitals, medications). To acquire the data needed to run performance estimates for Asthma-01, for example, it is necessary to have:

```{r asthma_01_tables, results="hide"}
data("nemsqar_patient_scene_table")
data("nemsqar_response_table")
data("nemsqar_situation_table")
data("nemsqar_medications_table")
data("nemsqar_vitals_table")
```

After loading, each object will appear in your R environment as a standard data frame.

## Inspecting Table Structure

Because R is new for many users, it is important to confirm that each dataset loaded correctly and contains the variables expected by the measure calculation functions. A few simple commands help verify structure:

### Review the structure of a loaded dataset
```{r glimpse_patient_scene_table}
# An easy to read quick look at the data using the `glimpse()` function from the famous `dplyr` package
dplyr::glimpse(nemsqar_patient_scene_table)
```

### Take a look at the first few rows of the dataset itself
```{r head_patient_scene_table}
# An abbreviated look at the actual data tables
head(nemsqar_patient_scene_table, n = 10)
```

These functions show variable names, data types, and the first few rows of the table. This step is critical, because each measure expects specific fields, and incorrect data types (for example, character instead of numeric) will cause functions to fail.

### Data types

Going forward, as these datasets were designed for this package, we will assume the data types of the columns are ready for us. However, in the wild, you will need to check the data types of your columns to ensure they meet `nemsqar` data validation requirements. If you do not, you will inevitably run into some friction by the way of error messages as `nemsqar` will validate your data.

### Dealing with problematic column names

Data coming out of EMS registries often have column names that are difficult to use when programming in R. Here, we will deal with this problem directly.

```{r clean_column_names, echo=TRUE, eval=TRUE}
# Define a custom cleaning function we can reuse
mr_clean_cols <- function(data) {
  # Clean column names using regex
  clean <- data |>
    dplyr::rename_with(
      .cols = tidyselect::everything(),
      ~ gsub(pattern = "\\.|\\(|-|\\s", replacement = "_", x = .)
    ) |>
    dplyr::rename_with(
      .cols = tidyselect::everything(),
      ~ gsub(pattern = "_+", replacement = "_", x = .)
    ) |>
    dplyr::rename_with(
      .cols = tidyselect::everything(),
      ~ gsub(pattern = "\\)", replacement = "", x = .)
    )

  # return cleaned columns
  return(clean)
}

# Clean column names using base functions and `dplyr`

# patient data
nemsqar_patient_scene_data <- nemsqar_patient_scene_table |>
  mr_clean_cols()

# response data
nemsqar_response_data <- nemsqar_response_table |>
  mr_clean_cols()

# situation data
nemsqar_situation_data <- nemsqar_situation_table |>
  mr_clean_cols()

# medications data
nemsqar_medications_data <- nemsqar_medications_table |>
  mr_clean_cols()

# vitals data
nemsqar_vitals_data <- nemsqar_vitals_table |>
  mr_clean_cols()

# `glimpse()` the cleaned columns of one dataset
dplyr::glimpse(nemsqar_patient_scene_data)
```

Now, special characters and whitespace are either removed or replaced with `_` so R can more easily recognize the column names, and we can avoid annoying conventions to find column names. We will need to do this again later so 

## Understanding Required Inputs
Each NEMSQA measure uses a defined set of input tables. `nemsqar` was designed to accept one single data source (via argument `df`) in case a user has their data structured in this way. Generally, though, this is not recommended. Instead, use the `*_table` arguments, where `*` is the name of a table, like `patient_scene_table`. It is best practice to store your data in different tables according to the NEMSIS structure such as having 1:1 *ePatient* and *eScene* data in one table, *eResponse* data in a separate table, *eSituation* data in another, and so on (i.e. via a Star Schema format). Although the EMS domain concepts may already be familiar, the R functions (and NEMSQA) require them to be organized in a consistent format. For example:

* **Asthma‑01** uses patient, response, situation, and medication tables.
* **Respiratory‑01** uses patient, response, situation, and vitals tables.
* **Safety‑01** uses patient and response tables.

The following sections demonstrate how to supply these inputs to `nemsqar` and how to run your first measure.

# Running the NEMSQA Measures Using `nemsqar`

With the required input tables loaded for *Asthma-01*, you are ready to calculate your first NEMSQA measure. Each measure in `nemsqar` is computed by a dedicated function. The applicable functions accept individual NEMSIS‑structured tables as arguments and return a standardized results dataset. 

## `nemsqar` workhorse functions
Each measure has two workhorse functions that do the heavy lifting for you. These functions follow a `name_##()` format. For example, *Asthma-01* is calculated by the `asthma_01()` function. In reality, `asthma_01()` is a wrapper function, as it calls another function `asthma_01_population()` which gathers the data and filters it down to the correct NEMSQA population for the measure. Then, `asthma_01()` calls a few internal helpers to calculate performance on the measure. Each NEMSQA measure has corresponding wrapper and population functions.

As an example, the code below demonstrates how to run **Asthma‑01** using the example data loaded in earlier sections. Remember, we loaded and inspected our data before. We use column names from each applicable table to fulfill each argument for fields required to calculate the given NEMSQA measure. 

```{r run_asthma_01, results="show"}
# Run `asthma_01` for a whole dataset, no grouping
asthma_01_all <- asthma_01(
  patient_scene_table = nemsqar_patient_scene_data,
  response_table = nemsqar_response_data,
  situation_table = nemsqar_situation_data,
  medications_table = nemsqar_medications_data,
  erecord_01_col = Incident_Patient_Care_Report_Number_PCR_eRecord_01,
  incident_date_col = Incident_Date,
  patient_DOB_col = Patient_Date_Of_Birth_ePatient_17,
  epatient_15_col = Patient_Age_ePatient_15,
  epatient_16_col = Patient_Age_Units_ePatient_16,
  eresponse_05_col = Response_Type_Of_Service_Requested_With_Code_eResponse_05,
  esituation_11_col = Situation_Provider_Primary_Impression_Code_And_Description_eSituation_11,
  esituation_12_col = Situation_Provider_Secondary_Impression_Description_And_Code_eSituation_12,
  emedications_03_col = Patient_Medication_Given_or_Administered_Description_And_RXCUI_Codes_List_eMedications_03,
  confidence_interval = T,
  method = "clopper-pearson",
  conf.level = 0.95
)

# print the results
asthma_01_all
```

Clearly, the ambulance service above needs to put a PI project into action right away! Asthmatic patients are not being taken care of appropriately. We may be curious to know how this service performs with specific types of patients.

```{r run_asthma_01_age, results="show"}
# Run `asthma_01` for a whole dataset, group by age units
asthma_01_age <- asthma_01(
  patient_scene_table = nemsqar_patient_scene_data,
  response_table = nemsqar_response_data,
  situation_table = nemsqar_situation_data,
  medications_table = nemsqar_medications_data,
  erecord_01_col = Incident_Patient_Care_Report_Number_PCR_eRecord_01,
  incident_date_col = Incident_Date,
  patient_DOB_col = Patient_Date_Of_Birth_ePatient_17,
  epatient_15_col = Patient_Age_ePatient_15,
  epatient_16_col = Patient_Age_Units_ePatient_16,
  eresponse_05_col = Response_Type_Of_Service_Requested_With_Code_eResponse_05,
  esituation_11_col = Situation_Provider_Primary_Impression_Code_And_Description_eSituation_11,
  esituation_12_col = Situation_Provider_Secondary_Impression_Description_And_Code_eSituation_12,
  emedications_03_col = Patient_Medication_Given_or_Administered_Description_And_RXCUI_Codes_List_eMedications_03,
  confidence_interval = TRUE,
  method = "clopper-pearson",
  conf.level = 0.95,
  .by = Patient_Age_Units_ePatient_16
)

# print the results
asthma_01_age
```

# Common Pitfalls for New R Users
Users who are new to R often encounter a few predictable issues when preparing data for measure calculation. The most common are listed below:


## Incorrect variable types
Many NEMSQA logic components require numeric fields. If values are read as character strings, the measure will fail. Use functions to manipulate data so it is in the correct format, as needed.


## Unintended name changes
Columns must match expected NEMSIS fields exactly. You can name your columns whatever you want, but the data from **eResponse.05** must go in the `eresponse_05_col` argument of any function to work. It is the nature of the data, not naming conventions, that will break your function calls in `nemsqar`.


## Missing required tables
Each measure requires a specific set of inputs. If one is missing, the function will return an error.


## Duplicated records
Measures assume that each patient or encounter corresponds to a single row in the input tables that are expected to be 1:1. Duplicate records can cause unexpected inclusion or exclusion, although `nemsqar` does try to help users with this out of the box. Always check your data and be aware of any Cartesian product that may be introduced unknowingly due to database issues or joins. 